Interactive Curvature Tensor Visualization on Digital Surfaces\thanks{This work has been mainly funded by \textsc{DigitalSnow} ANR-11-BS02-009, \textsc{KIDICO} ANR-2010-BLAN-0205 and \textsc{PRIMES Labex} ANR-11-LABX-0063/ ANR-11-IDEX-0007 research grants.}

### 2015-10-23 latex/article.tex
article

\sloppy In our context, we consider digital shapes (any subset $Z$ of $\Z^d$)
and boundaries of digital shapes $\DigBoundary{Z}$. We denote by $\DSh$ the
Gauss digitization of $\Shape$ in a $d-$dimensional grid with grid step $h$,
\emph{i.e.}  $\DSh \EqDef \{ \vz \in \Z^d,\, h\cdot\vz\in\Shape\}$. For such
digitized set $\DigShape\EqDef\DSh$, $\Body{\DigShape}{h}$ is a subset of $\R^d$
corresponding to the union of hypercubes centered at $h\cdot \vz$ for $\vz\in Z$
with edge length $h$. By doing so, both $\partial\Shape$ and
$\partial\Body{Z}{h}$ are topological boundary of objects lying in the same
space (See Fig. \ref{fig:notations}-b). Note that the combinatorial digital
boundary $\DigBoundary{Z}$ of $Z$ made of cells in a cellular Cartesian complex
(\emph{pointels}, \emph{linels}, \emph{surfels}, \ldots), can be trivially
embedded into $\R^d$ such that it coincides with $\partial\Body{Z}{h}$.

\begin{figure}[t]{\small
    \begin{center}
      \subfigure[]{\begin{overpic}[width=5cm]{figs/integral2D_2}
          \put(42,28){$\times$}
          \put(46,32){$\vx$}
          \put(55,45){$\Ball{R}{\vx}$}
          \put(1,10){$\dS$}
          \put(15,5){$\Shape$}
      \end{overpic}}~~~~~~
      \subfigure[]{\begin{overpic}[width=5cm]{figs/notation_2}
          \put(62,0.5){$h \cdot \vz \in \Shape \cap \Ball{R}{\vx}$}
          \put(55,45){$\Ball{R}{\vx}$}
          \put(42,28.5){$\times$}
          \put(45,32.5){$\vx$}
          \put(45,16.5){$h$}
          %\put(88,39){$\Proj(p)$}
          \put(1.5,10){$\partial \Shape$}
          \put(15,7){$\partial \Body{\DSh}{h}$}
      \end{overpic}}
      \vspace{-0.4cm}
    \end{center}}
    \caption{Integral invariant computation $(a)$ and
      notations $(b)$ in dimension 2 \cite{CVIU2014}.\label{fig:notations}}
\end{figure}

@@

\begin{figure}[!htbp]
  \begin{center}
    \subfigure{}{\includegraphics[width=0.9\textwidth]{figs/morton2}}
    %%\subfigure{}{\includegraphics[height=0.17\textheight]{figs/partitioning}}
    %%\subfigure{}{\includegraphics[height=0.17\textheight]{figs/quadtree_activefront}}
  \end{center}
\caption{Morton codes associated to cells of a linear quatree. Each
   morton code of a child cell is obtained by adding a suffix to its
  parent code \emph{(left)}. The adaptive representation consists of
  quadtree cells whose depth is view point dependent
  \emph{(middle)}. Finally, adaptive Marching Cubes is used to
  generate the triangulation \emph{(right)}.}
\label{fig_quadtree_partitionning}
\end{figure}

@@

Efficient spatial tree encoding can be
achieved using pointerless structures such as linear quadtrees or octrees ( see
for instance \textsc{Gargantini}~\cite{gargantini1982effective} ). This structure
%consists in indexing each cell by a \emph{Morton code}: the code of children
indexes each cell by a \emph{Morton code}: the code of children
cells are defined by the code of the parent suffixed by two bits (in dimension
2, three bits in dimension 3) (see Figure
\ref{fig_quadtree_partitionning}-\emph{left}). A cell's code encodes its
position \wrt its parent cell and its complete path to the tree root.
%Hence, whatever their depth, all cell codes are simply stored in a linear vector on memory.
Hence, the tree is fully represented as a linear vector of its leaves, whatever their depth.
Furthermore, cell operations such as subdivision, merging, neighbors
access can be efficiently implemented using bitwise operations on the Morton
code. In the following, we use the GPU implementation proposed by \textsc{Dupuy}
\textit{et al.}~\cite{dupuy2014quadtrees} that we extended to the third dimension.
By using their approach, Morton codes
and bitwise operations are optimized to match with GPU hardware requirements.

@@

As illustrated in Figure~\ref{fig_quadtree_partitionning}-\emph{middle}, we
propose a viewpoint dependent criterion to decide if a cell needs to be refined:
the closer we are to the camera, the finer the cells are.
%In addition to the
%distance criterion, we define an angular criterion to prevent refining cells
%outside the camera frustrum.
Using a criterion based on the distance between a
cell and the camera is well suited to GPU since it can be evaluated
independently on all cells. Hence, we can efficiently determine if a cell must
be subdivided or not. Figure~\ref{fig_lod_octree} illustrates the level of
details (LoD for short) criterion we use. More precisely, in dimension 2, if
$\alpha$ denotes the viewing angle, an object at distance $d$ from the camera
with diameter greater than $2\cdot d\cdot\tan(\alpha)$ is projected into more
than one pixel (see Figure~\ref{fig_lod_octree}-\emph{left}). Hence, the
distance criterion is based on the ratio (\emph{visibility ratio} in the
following) between the cell diameter $l(c)$ (power of 2 depending on the depth),
and $2\cdot d_c\cdot\tan(\alpha)$, $d_c$ being the distance between the cell
center and the camera. For a given cell $c$, split and merge decision are based
on this visibility ratio:
%
\begin{itemize}
  \item $c$ is split if its children
cells have a visibility ratio greater than  constant $k$\,;
  \item $c$ and its sibling cells are merged if their parent cell $c'$ has a visibility ratio lower than $k$\;
  \item otherwise, the cell $c$ stays for the next frame.
\end{itemize}
%
\begin{figure}[!h]
  \begin{center}
    \begin{overpic}[width=3.5cm]{figs/criterion}
      \put(33,40){$\alpha$}
      \put(50,45){$d_c$}
      \put(95,65){$l(c)$}
    \end{overpic}~~~~~
    \includegraphics[width=3.5cm]{viewlod2_small}~~~~~
    \includegraphics[width=3cm]{subdivision}
  \end{center}
  \caption{Notations \emph{(left)} and adaptive meshing in dimension 2 using the LoD distance and angular criterion \emph{(middle)}. Based on LoD criterion
    evaluation, the cell buffer is updated between each frame \emph{(right)}.}
  \label{fig_lod_octree}
\end{figure}
%
In addition to the distance criterion, we add an angular term in order
to prevent subdivision of cells far from the camera frustrum (see
Figure~\ref{fig_lod_octree}-\emph{middle}). We do not go into the
details but please note that stay, split and merge decisions of all cells are computed in
parallel from all cell morton codes.

@@

\begin{figure}
  \begin{center}
    {\includegraphics[width=4cm]{figs/approx1}}
    {\includegraphics[width=4cm]{figs/approx-reg-1}}
    {\includegraphics[width=4cm]{figs/approx-reg-2}}
  \end{center}
  \caption{Integral computation with an area estimation (\emph{left}) from pixel
    counting (\emph{middle}) or from densities (represented by different
    gray values) in a mipmap approximation of $\Shape$ (\emph{right}).}
  \label{fig:approx}
\end{figure}

@@

\begin{figure}
  \begin{center}
    {\includegraphics[width=4cm]{figs/approx-bh}}
    {\includegraphics[width=4cm]{figs/approx-ada}}
  \end{center}
  \caption{Adaptive approximation of the integration using a
    hierarchical decomposition of the ball $\Ball{R}{\vx}$ (\emph{left}), or a decomposition of $\Ball{R}{\vx}\cap\Shape$ (\emph{right}).}
  \label{fig:approx2}
\end{figure}

@@

%\todo[inline]{phrase sur approx si on met}

@@

\begin{figure}
  \begin{center}
    {\includegraphics[width=2.2cm]{figs/Octa_mean}}
    {\includegraphics[width=2.2cm]{figs/Octa_gaussian}}
    % {\includegraphics[width=2.7cm]{figs/Octa_k1}}
    % {\includegraphics[width=2.7cm]{figs/Octa_k2}}\\
    % {\includegraphics[width=2.7cm]{figs/Octa_dir_min}}
    % {\includegraphics[width=2.7cm]{figs/Octa_dir_max}}
    {\includegraphics[height=2.0cm]{figs/Octa_dir_min_zoom}}
    {\includegraphics[height=2.0cm]{figs/Octa_dir_max_zoom}}\\
    % {\includegraphics[height=4.0cm]{figs/EG_dragon_normal_zoom}}
    {\includegraphics[width=2.2cm]{figs/function_mean_0}}
    {\includegraphics[width=2.2cm]{figs/function_mean_1}}
    {\includegraphics[width=2.2cm]{figs/function_mean_2}}
    {\includegraphics[width=2.2cm]{figs/function_mean_3}}
    {\includegraphics[width=2.2cm]{figs/function_mean_4}}
  \end{center}
  \caption{\emph{First row}: Mean and Gaussian curvature estimation, (zoom of) first and second principal directions estimation on an OctaFlower with a digital domain of $130^3$.
  % \emph{Second row:} (Zoom of) First and second principal directions estimation on an OctaFlower.
  \emph{Third row:} Mean curvature computed in real-time on a dynamic object.}
  \label{fig:full}
\end{figure}

\begin{figure}
  \begin{center}
    \includegraphics[width=0.5\textwidth]{figs/graph_cmp_l0_l4_mean_octa130_Linf}
  \end{center}
  \caption{Error $L\infty$ on the measured Mean curvature between a $l_n$ approximation and the
  $l_0$ measurement. The red curve present the results when computing the volume from a mipmap quantitized to 0 or 1, and the green one
  present the results when computing it from the real mipmap values.}
  \label{fig:graph_convergence}
\end{figure}

@@

Then, we compare the approximation made by computing curvature from level $l \ge
l_0$ in Figure~\ref{fig:refine}. We can see that results using approximation
seems to fastly converge to the ground truth results.
Table~\ref{tab:full-res-stat} shows numerical results of $L_\infty$ error (worst
case) and $L_2$ error (mean), as well the number of \emph{mipmap texture}
fetches. We can first see that the number of texel fetch required to compute the
curvature reduces drastically when computing the approximations. This is
mandatory if we want to increase the framerate.

 in order to accelerate visualisation with the ground truth result obtained
at level $l_0$. We also compare them with the hierarchical algorithm discussed
in Section~\ref{sec:hierarchical}. The results are presented in
Table~\ref{tab:full-res-stat}, and illustrated in Figure~\ref{fig:refine}.

However, we can also note that the error increases quite slowly.
The errors obtained comparing $l_1$, or for bigger radii $l_2$ approximations
with the $l_0$ level are small. The errors measured comparing higher levels
are of course much bigger and those level are thus never used in practice.
They are presented here to illustrate how the refining process converges to
the $l_0$ level.

This convergence is shown in Figure \ref{fig:convergence}. It has been evaluated using
two different strategy for volume computation. The first one reads the interpolated
data from the mipmap is adds 1 to the volume if this value is above a threshold.
The second strategy adds directly to the volume the interpolated data read from the mipmap.
The results are shown Figure \ref{fig:graph_convergence}. It can be seen that the
interpolation done by the mipmap allows the resulting curvature to converge
faster to the $l_0$ level.

Thanks to this low error, higher approximation levels are visually
very similar to the $l_0$ level. We can therefore use them for exploring our
datasets in real time, changing the radii on the fly and visualising dynamic datasets.

When computing results using the hierachical algorithm, a higher error can be noted.
This is due to the precomputation of the subdivision that no longer ensures to fetch data
at the center of a mipmap cell. The interpolation done by the GPU reduces the bias but
does not remove it.

@@

      \multirow{2}{2.1cm}{$L_2$ error (\wrt $l_0$)} & $R=8$ & 3.51e-05 & & 2.67e-04 & 7.57e-05 & 2.02e-05 & 0\\
                               & $R=16$ & 4.43e-05 & 1.39e-04 & 4.16e-05 & 1.57e-05 & 2.07e-06 & 0\\
      \bottomrule
    \end{tabular}
  \end{center}
  %
  \caption{Comparison of number of texture fetches, $L_2$ and $L_\infty$ error
  obtained on OctaFlower with a digital domain of $130^3$ when computing mean
  curvature with two radii: 8 and 16, with hierarchical algorithm (H) and $l \ge
  l_0$ approximation algorithms. The object is triangulated at full resolution
  with a regular grid and contains 282,396 vertices.
  %
  \label{tab:full-res-stat}}
  %
\end{table}

\begin{figure}[!htbp]
  \begin{center}
   {\includegraphics[width=2.7cm]{figs/octa_r10_l3_a}}
   {\includegraphics[width=2.7cm]{figs/octa_r10_l2_a}}
   {\includegraphics[width=2.7cm]{figs/octa_r10_l1_a}}
   {\includegraphics[width=2.7cm]{figs/octa_r10_l0_a}}
  \end{center}
  \caption{Illustration of mean curvature computation on an OctaFlower (digital domain of $130^3$) using mipmap approximation with different levels : $l_3$, $l_2$, $l_1$ and $l_0$ (\ie no approximation).
  %
  For all illustrations, we use directly normal vectors computed by algorithm discussed in Section~\ref{sec:curv-tens-estim}}
  \label{fig:refine}
\end{figure}

@@

\begin{figure}[!htbp]
  \begin{center}
    \begin{tikzpicture}[scale=0.7]
      \begin{axis}[
          ybar stacked,
          ymode = log,
          ymax = 6000,
	  	  bar width=20pt,
	  	  nodes near coords,
	  	  every node near coord/.style={ font=\tiny, inner sep=0pt},
          enlargelimits=0.15,
          ylabel={\#Time (in ms)},
          symbolic x coords={Octa-R8, Octa-R16, Snow-R10, Snow-R20,
	      Dragon-R8, Dragon-R20},
          xtick=data,
          point meta=rawy,
          x tick label style={rotate=45,anchor=east},
        ]
        %Geometry Generation
        \addplot+[ybar, color=black, fill=cgeometry] plot coordinates {(Octa-R8,17.283000) (Octa-R16,18.458000)
          (Snow-R10,9.831000) (Snow-R20,9.653000)  (Dragon-R8,136.300000) (Dragon-R20,145.450000)};
        %l4
        \addplot+[ybar, color=black, fill=clquatre] plot coordinates {(Octa-R8,1) (Octa-R16,2.509000)
          (Snow-R10,6.085000) (Snow-R20,9.445000)  (Dragon-R8,3.389000) (Dragon-R20,20.228000)};
        %l3
        \addplot+[ybar, color=black, fill=cltrois] plot coordinates {(Octa-R8,2.527000) (Octa-R16,4.159000)
          (Snow-R10,10.629000) (Snow-R20,11.542000) (Dragon-R8,20.397000) (Dragon-R20,26.710000)};
        %l2
        \addplot+[ybar, color=black, fill=cldeux] plot coordinates {(Octa-R8,3.783000) (Octa-R16,9.006000)
          (Snow-R10,13.777000) (Snow-R20,37.700000) (Dragon-R8,29.967000) (Dragon-R20,92.435000)};
        %l1
        \addplot+[ybar, color=black, fill=clun] plot coordinates {(Octa-R8, 11.004000) (Octa-R16,38.687000)
          (Snow-R10,39.627000) (Snow-R20,217.954000) (Dragon-R8,59.840000) (Dragon-R20,506.549000)};
        %l0
        \addplot+[ybar, color=black, fill=clzero] plot coordinates {(Octa-R8, 37.273000) (Octa-R16,244.353000)
          (Snow-R10,236.194000) (Snow-R20,1655.680000) (Dragon-R8,322.674000) (Dragon-R20,3763.422000)};
        %\legend{\strut Geometry, \strut Curvature, \strut xx, \strut xx}
      \end{axis}
	  \end{tikzpicture}
%
    \begin{tikzpicture}[scale=0.7]
      \begin{axis}[
          ybar stacked,
          ymax = 6000,
          ymode = log,
	  	  bar width=20pt,
	  	  nodes near coords,
	  	  every node near coord/.style={ at={(0,0)}, font=\tiny, inner sep=0pt},
          enlargelimits=0.15,
          %ylabel={\#Time (in ms)},
          symbolic x coords={Octa-R8, Octa-R16, Snow-R10, Snow-R20,
	      Dragon-R8, Dragon-R20},
          xtick=data,
          point meta=rawy,
          x tick label style={rotate=45,anchor=east},
        ]
        %Geometry Generation
        \addplot+[ybar,color=black, fill=cgeometry] plot coordinates {(Octa-R8,17.283000) (Octa-R16,17.183000)
          (Snow-R10,9.383000) (Snow-R20,9.767000)  (Dragon-R8,127.347000) (Dragon-R20,127.166000)};
        %Hierarchical measurement
        \addplot+[ybar, color=black, fill=clzero] plot coordinates {(Octa-R8,18.095000) (Octa-R16,64.186000)
          (Snow-R10,96.747000) (Snow-R20,364.547000)  (Dragon-R8,147.662000) (Dragon-R20,787.459000)};
        %\legend{\strut Geometry, \strut Curvature, \strut xx, \strut xx}
      \end{axis}
     \end{tikzpicture}
     \end{center}
  %
  \caption{Timings in milliseconds (in logscale) obtained while visualizing an adaptive
  triangulation on three objects -- OctaFlower (digital domain of $130^3$ ), Snow
  Microstructures ($233^3$) and XYZ-Dragon ($510^3$) --  by computing the
  curvature with a regular grid (\emph{left}) and with hierarchical
  algorithm (\emph{right}), with two different radii for each object.
  %
  In \emph{orange color}: time required to extract the triangulation.
  %
  In \emph{blue color}: time required to compute the curvature tensor at
  different levels: from $l_4$ (light blue) to $l_0$ (dark blue).
  %
  Timings are given using a NVIDIA GeForce GTX 850M GPU.
  }
  %
  \label{fig:timings}
\end{figure}

@@

\begin{figure}
  \begin{center}
    {\includegraphics[height=4.0cm]{figs/xyzrgb_dragon-510_R8_mean}}
    {\includegraphics[height=3.5cm]{figs/snow_I08iso_233_r20_l0_m3}}
  \end{center}
  \caption{Illustration fig illustration complete (adaptative, ...) plusieurs
    objets.. + dynamique.
    Illustration of mean curvature computation of snow microstructures.
    %
    \emph{Third row:} mean curvature computed in real-time on a dynamic object.}
  \label{fig:adaptive}
\end{figure}

>>>>

\sloppy In our context, we consider digital shapes (any subset $Z$ of $\Z^d$)
and boundaries of digital shapes $\DigBoundary{Z}$. We denote by $\DSh$ the
Gauss digitization of $\Shape$ in a $d-$dimensional grid with grid step $h$,
\emph{i.e.}  $\DSh \EqDef \{ \vz \in \Z^d,\, h\cdot\vz\in\Shape\}$. For such
digitized set $\DigShape\EqDef\DSh$, $\Body{\DigShape}{h}$ is a subset of $\R^d$
corresponding to the union of hypercubes centered at $h\cdot \vz$ for $\vz\in Z$
with edge length $h$. By doing so, both $\partial\Shape$ and
$\partial\Body{Z}{h}$ are topological boundary of objects lying in the same
space (see Fig. \ref{fig:notations}-b). Note that the combinatorial digital
boundary $\DigBoundary{Z}$ of $Z$ made of cells in a cellular Cartesian complex
(\emph{pointels}, \emph{linels}, \emph{surfels}, \ldots), can be trivially
embedded into $\R^d$ such that it coincides with $\partial\Body{Z}{h}$.

\begin{figure}[t]{\small
    \begin{center}
      {\begin{overpic}[width=5cm]{figs/integral2D_2}
          \put(42,28){$\times$}
          \put(46,32){$\vx$}
          \put(55,45){$\Ball{R}{\vx}$}
          \put(1,10){$\dS$}
          \put(15,5){$\Shape$}
      \end{overpic}}~~~~~~
      {\begin{overpic}[width=5cm]{figs/notation_2}
          \put(62,0.5){$h \cdot \vz \in \Shape \cap \Ball{R}{\vx}$}
          \put(55,45){$\Ball{R}{\vx}$}
          \put(42,28.5){$\times$}
          \put(45,32.5){$\vx$}
          \put(45,16.5){$h$}
          %\put(88,39){$\Proj(p)$}
          \put(1.5,10){$\partial \Shape$}
          \put(15,7){$\partial \Body{\DSh}{h}$}
      \end{overpic}}
      \vspace{-0.4cm}
    \end{center}}
    \caption{Integral invariant computation (\emph{left}) and
      notations (\emph{right}) in dimension 2 \cite{CVIU2014}.\label{fig:notations}}
    \vspace{-0.35cm}
\end{figure}

@@

\begin{figure}[!htbp]
  \vspace{-0.35cm}
  \begin{center}
    \subfigure{}{\includegraphics[width=0.9\textwidth]{figs/morton2}}
    %%\subfigure{}{\includegraphics[height=0.17\textheight]{figs/partitioning}}
    %%\subfigure{}{\includegraphics[height=0.17\textheight]{figs/quadtree_activefront}}
  \end{center}
  \vspace{-0.35cm}
\caption{Morton codes associated to cells of a linear quatree. Each
   morton code of a child cell is obtained by adding a suffix to its
  parent code \emph{(left)}. The adaptive representation consists of
  quadtree cells whose depth is view point dependent
  \emph{(middle)}. Finally, adaptive Marching Cubes is used to
  generate the triangulation \emph{(right)}.}
\label{fig_quadtree_partitionning}
\end{figure}

@@

Efficient spatial tree encoding can be
achieved using pointerless structures such as linear quadtrees or octrees (see
for instance \textsc{Gargantini}~\cite{gargantini1982effective}). This structure
%consists in indexing each cell by a \emph{Morton code}: the code of children
indexes each cell by a \emph{Morton code}: the code of children
cells are defined by the code of the parent suffixed by two bits (in dimension
2, three bits in dimension 3) (see Figure
\ref{fig_quadtree_partitionning}-\emph{left}). A cell's code encodes its
position \wrt its parent cell and its complete path to the tree root.
%Hence, whatever their depth, all cell codes are simply stored in a linear vector on memory.
Hence, the tree is fully represented as a linear vector of its leaves, whatever their depth.
Furthermore, cell operations such as subdivision, merging, neighbors
access can be efficiently implemented using bitwise operations on the Morton
code. In the following, we use the GPU implementation proposed by \textsc{Dupuy}
\textit{et al.}~\cite{dupuy2014quadtrees} that we extended to the third dimension.
By using their approach, Morton codes
and bitwise operations are optimized to match with GPU hardware requirements.

@@

As illustrated in Figure~\ref{fig_quadtree_partitionning}-\emph{middle}, we
propose a viewpoint dependent criterion to decide if a cell needs to be refined:
the closer we are to the camera, the finer the cells are.
%In addition to the
%distance criterion, we define an angular criterion to prevent refining cells
%outside the camera frustrum.
Using a criterion based on the distance between a
cell and the camera is well suited to GPU since it can be evaluated
independently on all cells. Hence, we can efficiently determine if a cell must
be subdivided or not. Figure~\ref{fig_lod_octree} illustrates the level of
details (LoD for short) criterion we use. More precisely, in dimension 2, if
$\alpha$ denotes the viewing angle, an object at distance $d$ from the camera
with diameter greater than $2\cdot d\cdot\tan(\alpha)$ is projected into more
than one pixel (see Figure~\ref{fig_lod_octree}-\emph{left}). Hence, the
distance criterion is based on the ratio (\emph{visibility ratio} in the
following) between the cell diameter $l(c)$ (power of 2 depending on the depth),
and $2\cdot d_c\cdot\tan(\alpha)$, $d_c$ being the distance between the cell
center and the camera. For a given cell $c$, split and merge decision are based
on this visibility ratio:
%
\begin{itemize}
  \item $c$ is split if its children
cells have a visibility ratio greater than  constant $k$\,;
  \item $c$ and its sibling cells are merged if their parent cell $c'$ has a visibility ratio lower than $k$\;
  \item otherwise, the cell $c$ stays for the next frame.
\end{itemize}
%
\begin{figure}[!h]
  \begin{center}
    \begin{overpic}[width=3.5cm]{figs/criterion}
      \put(33,40){$\alpha$}
      \put(50,45){$d_c$}
      \put(95,65){$l(c)$}
    \end{overpic}~~~~~
    \includegraphics[width=3cm]{viewlod2_small}~~~~~
    \includegraphics[width=3cm]{subdivision}
  \end{center}
  \caption{Notations \emph{(left)} and adaptive meshing in dimension 2 using the LoD distance and angular criterion \emph{(middle)}. Based on LoD criterion
    evaluation, the cell buffer is updated between each frame \emph{(right)}.}
  \label{fig_lod_octree}
\end{figure}
%
In addition to the distance criterion, we add an angular term in order
to prevent subdivision of cells far from the camera frustrum (see
Figure~\ref{fig_lod_octree}-\emph{middle}). We do not go into the
details but please note that stay, split and merge decisions of all cells are computed in
parallel from all cell morton codes.

@@

\begin{figure}
  \vspace{-0.35cm}
  \begin{center}
    {\includegraphics[width=4cm]{figs/approx1}}
    {\includegraphics[width=4cm]{figs/approx-reg-1}}
    {\includegraphics[width=4cm]{figs/approx-reg-2}}
  \end{center}
  \vspace{-0.35cm}
  \caption{Integral computation with an area estimation (\emph{left}) from pixel
    counting (\emph{middle}) or from densities (represented by different
    gray values) in a mipmap approximation of $\Shape$ (\emph{right}).}
  \label{fig:approx}
  \vspace{-0.35cm}
\end{figure}

@@

\begin{figure}
  \vspace{-0.35cm}
  \begin{center}
    {\includegraphics[width=4cm]{figs/approx-bh}}
    {\includegraphics[width=4cm]{figs/approx-ada}}
  \end{center}
  \vspace{-0.35cm}
  \caption{Adaptive approximation of the integration using a
    hierarchical decomposition of the ball $\Ball{R}{\vx}$ (\emph{left}), or a decomposition of $\Ball{R}{\vx}\cap\Shape$ (\emph{right}).}
  \label{fig:approx2}
\end{figure}

@@

@@

\begin{figure}
  \vspace{-0.35cm}
  \begin{center}
    {\includegraphics[width=2.2cm]{figs/Octa_mean}}
    {\includegraphics[width=2.2cm]{figs/Octa_gaussian}}
    % {\includegraphics[width=2.7cm]{figs/Octa_k1}}
    % {\includegraphics[width=2.7cm]{figs/Octa_k2}}\\
    % {\includegraphics[width=2.7cm]{figs/Octa_dir_min}}
    % {\includegraphics[width=2.7cm]{figs/Octa_dir_max}}
    {\includegraphics[height=2.0cm]{figs/Octa_dir_min_zoom}}
    {\includegraphics[height=2.0cm]{figs/Octa_dir_max_zoom}}\\
    % {\includegraphics[height=4.0cm]{figs/EG_dragon_normal_zoom}}
    {\includegraphics[width=2.2cm]{figs/function_mean_0}}
    {\includegraphics[width=2.2cm]{figs/function_mean_1}}
    {\includegraphics[width=2.2cm]{figs/function_mean_2}}
    {\includegraphics[width=2.2cm]{figs/function_mean_3}}
    {\includegraphics[width=2.2cm]{figs/function_mean_4}}
  \end{center}
  \vspace{-0.35cm}
  \caption{\emph{First row}: Mean and Gaussian curvature estimation, (zoom of) first and second principal directions estimation on an OctaFlower with a digital domain of $130^3$.
  % \emph{Second row:} (Zoom of) First and second principal directions estimation on an OctaFlower.
  \emph{Third row:} Mean curvature computed in real-time on a dynamic object.}
  \label{fig:full}
\end{figure}

% \begin{figure}
%   \begin{center}
%     \includegraphics[width=0.5\textwidth]{figs/graph_cmp_l0_l4_mean_octa130_Linf}
%   \end{center}
%   \caption{Error $L\infty$ on the measured Mean curvature between a $l_n$ approximation and the
%   $l_0$ measurement. The red curve present the results when computing the volume from a mipmap quantitized to 0 or 1, and the green one
%   present the results when computing it from the real mipmap values.}
%   \label{fig:graph_convergence}
% \end{figure}

@@

Then, we compare the approximation made by computing curvature from level $l \ge
l_0$ in Figure~\ref{fig:refine}. We can see that results using approximation
seems to fastly converge to the ground truth results.
Table~\ref{tab:full-res-stat} shows numerical results of $L_\infty$ error (worst
case) and $L_2$ error (mean), as well the number of \emph{mipmap texture}
fetches. We can also see that the number of texel fetch, required to compute the
curvature, reduces drastically when computing approximations. When measuring
results with higher levels, we get consequently much bigger errors. Those level
are thus never used in practice, they are presented here to illustrate how the
refining process converges to the $l_0$ level. We see that levels close to $l_0$
give few errors than $l_0$, and should be considered for exploring our datasets
in real time, changing parameters on-the-fly and visualizing dynamic datasets.

We also compare them with the hierarchical algorithm (as discussed in
Section~\ref{sec:hierarchical}). This method introduce a higher error than
$l_0$. This is due to the precomputation of the subdivision that no longer
ensures to fetch data at the center of a mipmap cell. Then, the GPU performes an
interpolation to reduce the bias, but it doesn't remove it.

@@

      \multirow{2}{2.1cm}{$L_2$ error (\wrt $l_0$)} & $R=8$ & 3.51e-05 & & 2.67e-04 & 7.57e-05 & 2.02e-05 & 0\\
                               & $R=16$ & 4.43e-05 & 1.39e-04 & 4.16e-05 & 1.57e-05 & 2.07e-06 & 0\\
      \bottomrule
    \end{tabular}
  \end{center}
  %
  \caption{Comparison of number of texture fetches, $L_2$ and $L_\infty$ error
  obtained on OctaFlower with a digital domain of $130^3$ when computing mean
  curvature with two radii: 8 and 16, with hierarchical algorithm (H) and $l \ge
  l_0$ approximation algorithms. The object is triangulated at full resolution
  with a regular grid and contains 282,396 vertices.
  %
  \label{tab:full-res-stat}}
  \vspace{-1.5cm}
  %
\end{table}

\begin{figure}[!htbp]
  \vspace{-0.35cm}
  \begin{center}
   {\includegraphics[width=2.7cm]{figs/octa_r10_l3_a}}
   {\includegraphics[width=2.7cm]{figs/octa_r10_l2_a}}
   {\includegraphics[width=2.7cm]{figs/octa_r10_l1_a}}
   {\includegraphics[width=2.7cm]{figs/octa_r10_l0_a}}
  \end{center}
  \vspace{-0.35cm}
  \caption{Illustration of mean curvature computation on an OctaFlower (digital domain of $130^3$) using mipmap approximation with different levels : $l_3$, $l_2$, $l_1$ and $l_0$ (\ie no approximation).
  %
  For all illustrations, we use directly normal vectors computed by algorithm discussed in Section~\ref{sec:curv-tens-estim}}
  \label{fig:refine}
  \vspace{-0.35cm}
\end{figure}

@@

\begin{figure}[!htbp]
  \begin{center}
    \begin{tikzpicture}[scale=0.7]
      \begin{axis}[
          ybar stacked,
          ymode = log,
          ymax = 6000,
	  	  bar width=20pt,
	  	  nodes near coords,
	  	  every node near coord/.style={ font=\tiny, inner sep=0pt},
          enlargelimits=0.15,
          ylabel={\#Time (in ms)},
          symbolic x coords={Octa-R8, Octa-R16, Snow-R10, Snow-R20,
	      Dragon-R8, Dragon-R20},
          xtick=data,
          point meta=rawy,
          x tick label style={rotate=45,anchor=east},
        ]
        %Geometry Generation
        \addplot+[ybar, color=black, fill=cgeometry] plot coordinates {(Octa-R8,17.283000) (Octa-R16,18.458000)
          (Snow-R10,9.831000) (Snow-R20,9.653000)  (Dragon-R8,136.300000) (Dragon-R20,145.450000)};
        %l4
        \addplot+[ybar, color=black, fill=clquatre] plot coordinates {(Octa-R8,1) (Octa-R16,2.509000)
          (Snow-R10,6.085000) (Snow-R20,9.445000)  (Dragon-R8,3.389000) (Dragon-R20,20.228000)};
        %l3
        \addplot+[ybar, color=black, fill=cltrois] plot coordinates {(Octa-R8,2.527000) (Octa-R16,4.159000)
          (Snow-R10,10.629000) (Snow-R20,11.542000) (Dragon-R8,20.397000) (Dragon-R20,26.710000)};
        %l2
        \addplot+[ybar, color=black, fill=cldeux] plot coordinates {(Octa-R8,3.783000) (Octa-R16,9.006000)
          (Snow-R10,13.777000) (Snow-R20,37.700000) (Dragon-R8,29.967000) (Dragon-R20,92.435000)};
        %l1
        \addplot+[ybar, color=black, fill=clun] plot coordinates {(Octa-R8, 11.004000) (Octa-R16,38.687000)
          (Snow-R10,39.627000) (Snow-R20,217.954000) (Dragon-R8,59.840000) (Dragon-R20,506.549000)};
        %l0
        \addplot+[ybar, color=black, fill=clzero] plot coordinates {(Octa-R8, 37.273000) (Octa-R16,244.353000)
          (Snow-R10,236.194000) (Snow-R20,1655.680000) (Dragon-R8,322.674000) (Dragon-R20,3763.422000)};
        %\legend{\strut Geometry, \strut Curvature, \strut xx, \strut xx}
      \end{axis}
	  \end{tikzpicture}
%
    \begin{tikzpicture}[scale=0.7]
      \begin{axis}[
          ybar stacked,
          ymax = 6000,
          ymode = log,
	  	  bar width=20pt,
	  	  nodes near coords,
	  	  every node near coord/.style={ at={(0,0)}, font=\tiny, inner sep=0pt},
          enlargelimits=0.15,
          %ylabel={\#Time (in ms)},
          symbolic x coords={Octa-R8, Octa-R16, Snow-R10, Snow-R20,
	      Dragon-R8, Dragon-R20},
          xtick=data,
          point meta=rawy,
          x tick label style={rotate=45,anchor=east},
        ]
        %Geometry Generation
        \addplot+[ybar,color=black, fill=cgeometry] plot coordinates {(Octa-R8,17.283000) (Octa-R16,17.183000)
          (Snow-R10,9.383000) (Snow-R20,9.767000)  (Dragon-R8,127.347000) (Dragon-R20,127.166000)};
        %Hierarchical measurement
        \addplot+[ybar, color=black, fill=clzero] plot coordinates {(Octa-R8,18.095000) (Octa-R16,64.186000)
          (Snow-R10,96.747000) (Snow-R20,364.547000)  (Dragon-R8,147.662000) (Dragon-R20,787.459000)};
        %\legend{\strut Geometry, \strut Curvature, \strut xx, \strut xx}
      \end{axis}
     \end{tikzpicture}
     \end{center}
  %
  \vspace{-0.35cm}
  \caption{Timings in milliseconds (in logscale) obtained while visualizing an adaptive
  triangulation on three objects -- OctaFlower (digital domain of $130^3$ ), Snow
  Microstructures ($233^3$) and XYZ-Dragon ($510^3$) --  by computing the
  curvature with a regular grid (\emph{left}) and with hierarchical
  algorithm (\emph{right}), with two different radii for each object.
  %
  In \emph{orange color}: time required to extract the triangulation.
  %
  In \emph{blue color}: time required to compute the curvature tensor at
  different levels: from $l_4$ (light blue) to $l_0$ (dark blue).
  %
  Timings are given using a NVIDIA GeForce GTX 850M GPU.
  }
  %
  \label{fig:timings}
  \vspace{-0.35cm}
\end{figure}

@@

\begin{figure}
  \vspace{-0.35cm}
  \begin{center}
    {\includegraphics[height=3.5cm]{figs/xyzrgb_dragon-510_R8_mean}}
    {\includegraphics[height=3.0cm]{figs/snow_I08iso_233_r20_l0_m3}}
  \end{center}
  \caption{Illustration of mean curvature computation of two large datasets: XYZ-Dragon (with digital domain of $512^3$) and Snow microstructures ($233^3$).}
  \label{fig:adaptive}
  \vspace{-0.35cm}
\end{figure}
